---
title: "final_project"
author: "Yucong Gao"
date: "4/22/2022"
output: github_document
---

```{r setup, include=FALSE}
library(psych)
library(GPArotation)
library(lavaan)
```

### Package Loading
```{r}
# Function: installing and loading of packages
install_load <- function (packages)  {   
   
  # Start loop to determine if each package is installed
   for(package in packages){

       # If package is installed locally, load
       if(package %in% rownames(installed.packages()))
          do.call('library', list(package))

       # If package is not installed locally, download, then load
       else {
          install.packages(package, dependencies = TRUE)
          do.call("library", list(package))
       }
   } 
}

# Generic libraries loading
libs <- c("ggplot2", "maps", "plotly", "plyr", "dplyr", "rworldmap","stringr","lubridate", "plotly", "reshape2", "magrittr", "ggthemes", "tidyr", "DT", "lubridate","RColorBrewer")
install_load(libs)

# Specific methods libraries loading
libs.methods <- c("C50", "lattice", "caret", "nnet", "e1071","Matrix", "foreach","glmnet","C50","randomForest","ipred","rpart")
install_load(libs.methods)

# Data loading
data <- read.csv("./survey.csv")
```

## Data Cleaning
```{r}
# To delete no important elements
data <- data[ , !(names(data) %in% "state")]
data <- data[ , !(names(data) %in% "Timestamp")]
data <- data[ , !(names(data) %in% "comments")]
data <- data[ , !(names(data) %in% "self_employed")]


# Gender unification.
data$Gender %<>% str_to_lower()

male_str <- c("male", "m", "male-ish", "maile", "mal", "male (cis)", "make", "male ", "man","msle", "mail", "malr","cis man", "cis male")
trans_str <- c("trans-female", "something kinda male?", "queer/she/they", "non-binary","nah", "all", "enby", "fluid", "genderqueer", "androgyne", "agender", "male leaning androgynous", "guy (-ish) ^_^", "trans woman", "neuter", "female (trans)", "queer", "ostensibly male, unsure what that really means" )
female_str <- c("cis female", "f", "female", "woman",  "femake", "female ","cis-female/femme", "female (cis)", "femail")

data$Gender <- sapply(as.vector(data$Gender), function(x) if(x %in% male_str) "male" else x )
data$Gender <- sapply(as.vector(data$Gender), function(x) if(x %in% female_str) "female" else x )
data$Gender <- sapply(as.vector(data$Gender), function(x) if(x %in% trans_str) "trans" else x )
data %<>% filter(Gender != "a little about you")
data %<>% filter(Gender != "guy (-ish) ^_^")
data %<>% filter(Gender != "p")


# Age categorization
data$Age<-cut(data$Age, c(-Inf,20,35,65,Inf))

# NA values detection and deleting the row.
sapply(data, function(x) sum(is.na(x)))
data <- data[!is.na(data$work_interfere),]

# Saving the original data with all importants variables 
data.origin <- data
```


### Convert data to categories
age: (-Inf,20] ~ 1, (20,35] ~ 2, (35,65] ~ 3, (65, Inf] ~ 4 <br>
gender: male ~ 1, female ~ 2, trans ~ 3 <br>
country: United State ~ 1, others ~ 0 <br>
family_history: yes ~ 1, no ~ 0 <br> 
work_inference: never ~ 1, rarely ~ 2, sometimes ~ 3, often ~ 4 <br>
no_employees: 1-5 ~ 1, 6-25 ~ 2, 26-100 ~ 3, 100-500 ~ 4, 500-1000 ~ 5, more than 1000 ~ 6 <br>
remote_work: yes ~ 1, no ~ 0 <br>
tech_company: yes ~ 1, no ~ 0 <br>
benefits: yes ~ 1, no ~ 2, don't know ~ 3 <br>
care_options: yes ~ 1, no ~ 2, TRUE ~ 3 <br>
wellness_program: yes ~ 1, no ~ 2, don't know ~ 3 <br>
seek_help: yes ~ 1, no ~ 2, don't know ~ 3 <br>
anonymity: yes ~ 1, no ~ 2, don't know ~ 3 <br> 
leave: Very easy ~ 1, Somewhat easy ~ 2, Somewhat difficult ~ 3, Very difficult ~ 4, Don't know ~ 5 <br>
mental_health_consequence: Yes ~ 1, No ~ 2, Maybe ~ 3 <br>
phys_health_consequence: Yes ~ 1, No ~ 2, Maybe ~ 3 <br>
coworkers: Yes ~ 1, No ~ 2, Some of them ~ 3 <br>
supervisor: Yes ~ 1, No ~ 2, Some of them ~ 3 <br>
mental_health_interview: Yes ~ 1, No ~ 2, Maybe ~ 3 <br>
phys_health_interview: Yes ~ 1, No ~ 2, Maybe ~ 3 <br>
mental_vs_physical: yes ~ 1, no ~ 2, don't know ~ 3 <br>
obs_consequence: Yes ~ 1, No ~ 0 <br>


```{r}
dat = tibble(data.origin) %>% janitor::clean_names()

dat_tibble = 
   dat %>% 
   mutate(age = factor(case_when(age == "(-Inf,20]" ~ 1, 
                                 age == "(20,35]" ~ 2, 
                                 age == "(35,65]" ~ 3, 
                                 age == "(65, Inf]" ~ 4)), 
          gender = factor(case_when(gender == "male" ~ 1, 
                                    gender == "female" ~ 2, 
                                    TRUE ~ 3)), 
          country = factor(case_when(country == "United States" ~ 1, 
                                     TRUE ~ 2)), 
          family_history = factor(case_when(family_history == "Yes" ~ 1, 
                                            TRUE ~ 2)), 
          work_interfere = factor(case_when(work_interfere == "Never" ~ 1, 
                                            work_interfere == "Rarely" ~ 2,
                                            work_interfere == "Sometimes" ~ 3,
                                            work_interfere == "Often" ~ 4)), 
          no_employees = factor(case_when(no_employees == "1-5" ~ 1, 
                                          no_employees == "6-25" ~ 2,
                                          no_employees == "26-100" ~ 3,
                                          no_employees == "100-500" ~ 4,
                                          no_employees == "500-1000" ~ 5,
                                          no_employees == "More than 1000" ~ 6)), 
          remote_work = factor(case_when(remote_work == "Yes" ~ 1, 
                                         TRUE ~ 2)), 
          tech_company = factor(case_when(tech_company == "Yes" ~ 1, 
                                         TRUE ~ 2)), 
          benefits = factor(case_when(benefits == "Yes" ~ 1, 
                                      benefits == "No" ~ 2,
                                      T ~ 3)), 
          care_options = factor(case_when(care_options == "Yes" ~ 1, 
                                          care_options == "No" ~ 2,
                                          T ~ 3)), 
          wellness_program = factor(case_when(wellness_program == "Yes" ~ 1, 
                                              wellness_program == "No" ~ 2,
                                              TRUE ~ 3)), 
          seek_help = factor(case_when(seek_help == "Yes" ~ 1, 
                                       seek_help == "No" ~ 2,
                                       TRUE ~ 3)), 
          anonymity = factor(case_when(anonymity == "Yes" ~ 1, 
                                       anonymity == "No" ~ 2,
                                       TRUE ~ 3)), 
          leave = factor(case_when(leave == "Very easy" ~ 1, 
                                   leave == "Somewhat easy" ~ 2,
                                   leave == "Somewhat difficult" ~ 3,
                                   leave == "Very difficult" ~ 4,
                                   TRUE ~ 5)), 
          
          mental_health_consequence = factor(case_when(mental_health_consequence == "Yes" ~ 1, 
                                                       mental_health_consequence == "No" ~ 2,
                                                       TRUE ~ 3)), 
          phys_health_consequence = factor(case_when(phys_health_consequence == "Yes" ~ 1, 
                                                     phys_health_consequence == "No" ~ 2,
                                                     TRUE ~ 3)),
          coworkers = factor(case_when(coworkers == "Yes" ~ 1, 
                                       coworkers == "No" ~ 2,
                                       TRUE ~ 3)),
          supervisor = factor(case_when(supervisor == "Yes" ~ 1, 
                                       supervisor == "No" ~ 2,
                                       TRUE ~ 3)),
          mental_health_interview = factor(case_when(mental_health_interview == "Yes" ~ 1, 
                                       mental_health_interview == "No" ~ 2,
                                       TRUE ~ 3)),
          phys_health_interview = factor(case_when(phys_health_interview == "Yes" ~ 1, 
                                       phys_health_interview == "No" ~ 2,
                                       TRUE ~ 3)),
          mental_vs_physical = factor(case_when(mental_vs_physical == "Yes" ~ 1, 
                                       mental_vs_physical == "No" ~ 2,
                                       TRUE ~ 3)), 
          obs_consequence = factor(case_when(obs_consequence == "Yes" ~ 1, 
                                             TRUE ~ 2)))





dat_df = as.data.frame(dat_tibble)
```


## PCA

```{r}
features = dat_df[, c(6:23)]
for (i in 1:18){
   features[,i] = as.numeric(features[,i])
}

```

demographic variables (age, gender, country, state, family history) are not included in the PCA
response variable - treatment also not included in the pca

Parallel analysis finds that 5 eigenvalue larger than the expected "random chance" eigenvalues based on randomly resampled data. So in the following exploratory factor analysis, we decide to explore 3, 4, 5, 6 factors to fit the model
```{r}

depressionprl = fa.parallel(features, cor = "poly", fa = "pc")

depressionprl$pc.values

depressionprl$pc.sim
```


## Exploratory Factor Analyais

res_5 with 5 latent factors have the lowest BIC, so we use this five factors for our CFA
```{r}
# data to use for efa
dep_efa = features
res_3 = fa(dep_efa, 3, cor = "poly", rotate = "geominQ", fm = "wls")
res_4 = fa(dep_efa, 4, cor = "poly", rotate = "geominQ", fm = "wls")
res_5 = fa(dep_efa, 5, cor = "poly", rotate = "geominQ", fm = "wls")
res_6 = fa(dep_efa, 6, cor = "poly", rotate = "geominQ", fm = "wls")

res_5$loadings


```

## Confirmatory Factor Analysis
```{r}

```

